Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-26 prop_iop.tex
prop_letter.tex \section{Gleason} Updating...

Previously, we proved that if $\rho'$ is consistent with $\iota\circ\mu_{\rho}^{\mathrm{B}}$,
then $\rho'$ is close to $\rho$. Conversely, if $\rho'$ is close
to $\rho$, $\rho'$ is consistent with $\iota\circ\mu_{\rho}^{\mathrm{B}}$.
Before we introduce the lemma, we start the following definition.

\begin{definition}A mixed state\nb$\rho$ is called totally mixed
if $0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$ for any $P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $.\end{definition}

For example, any pure state is not totally mixed, and $\mathbb{1}/3=\left(\proj{0}+\proj{1}+\proj{2}\right)/3$
is a totally mixed over a three-dimensional Hilbert space.

\begin{lemma}Given a $d$-dimensional Hilbert space with the set
of all projectors\nb$\events$, and an interval-valued function\nb$\iota:\left[0,1\right]\rightarrow\mathscr{I}$.
For two mixed state\nb$\rho$ and $\rho'$, if the following conditions
holds.
\begin{itemize}
\item If $\left|x-y\right|\le\delta$, we must have $y\in\iota\left(x\right)$.
\item $\left\Vert \rho-\rho'\right\Vert \le\delta/d$.
\item $\rho$ is a totally mixed.
\end{itemize}
Then, $\rho'\in\coreBorn\left(\iota\circ\mu_{\rho}^{\mathrm{B}},\events\right)$.\end{lemma}

\begin{proof}Since
\begin{equation}
\delta/d\ge\left\Vert \rho'-\rho\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{\rho'-\rho}{\psi}\right|\,,
\end{equation}
we have
\begin{equation}
\delta/d\ge\left|\melem{\psi}{\rho'-\rho}{\psi}\right|=\left|\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)\right|\label{eq:Born_rule_difference}
\end{equation}
for all normalized pure state\nb$\ket{\psi}$.

When $P\in\left\{ \mathbb{0},\mathbb{1}\right\} $, it is obvious
$\mu_{\rho'}^{\mathrm{B}}\left(P\right)\in\iota\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$.

When $P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $,
there exists an orthonormal set $\left\{ \ket{i}\right\} _{i=0}^{d'-1}$
such that $P=\sum_{i=0}^{d'-1}\proj{i}$. By applying equation~(\ref{eq:Born_rule_difference}),
\begin{equation}
\left|\mu_{\rho'}^{\mathrm{B}}\left(P\right)-\mu_{\rho}^{\mathrm{B}}\left(P\right)\right|\le\sum_{i=0}^{d'-1}\left|\mu_{\rho'}^{\mathrm{B}}\left(\proj{i}\right)-\mu_{\rho}^{\mathrm{B}}\left(\proj{i}\right)\right|\le\delta\,.
\end{equation}

Since $P$ is neither $\mathbb{0}$ nor $\mathbb{1}$, $\mu_{\rho}^{\mathrm{B}}\left(P\right)\in\left(0,1\right)$.
Hence, we have $\mu_{\rho'}^{\mathrm{B}}\left(P\right)\in\iota\left(\mu_{\rho}^{\mathrm{B}}\left(P\right)\right)$.\end{proof}

The previous two lemmas can be applied on example~\ref{ex:equilateral},
we can estimate the states consistent with a QIVPM more tightly.

\begin{example}Given the additive interval-valued function\nb$\widehat{\iota}_{n}:\left[0,1\right]\rightarrow\widehat{\mathscr{I}}_{n}$,
the maximum length of intervals in $\widehat{\mathscr{I}}_{n}$ is
$1/n$, and If $\left|x-y\right|\le\left(4n\right)^{-1}$, we must
have $y\in\widehat{\iota}_{n}\left(x\right)$. Therefore, for any
totally mixed state\nb$\rho$ on a $d$-dimensional Hilbert space,
we have
\begin{itemize}
\item If $\left\Vert \rho-\rho'\right\Vert \le1/\left(4dn\right)$, then
$\rho'\in\coreBorn\left(\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)$.
\item If $\rho'\in\coreBorn\left(\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)$,
then $\left\Vert \rho-\rho'\right\Vert \le1/n$.
\end{itemize}
\end{example}

>>>>

The previous lemma can be applied on example~\ref{ex:equilateral},
we can estimate the states consistent with a QIVPM more tightly.

\begin{example}Given the additive interval-valued function\nb$\widehat{\iota}_{n}:\left[0,1\right]\rightarrow\widehat{\mathscr{I}}_{n}$,
the maximum length of intervals in $\widehat{\mathscr{I}}_{n}$ is
$1/n$. Therefore, for any totally mixed state\nb$\rho$ on a $d$-dimensional
Hilbert space, if $\rho'\in\coreBorn\left(\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)$,
then $\left\Vert \rho-\rho'\right\Vert \le1/n$.\end{example}

Previously, we proved that if $\rho'$ is consistent with $\iota\circ\mu_{\rho}^{\mathrm{B}}$,
then $\rho'$ is close to $\rho$. However, if $\rho'$ is close to
$\rho$, there is no way to guarantee $\rho'$ is consistent with
$\iota\circ\mu_{\rho}^{\mathrm{B}}$ because no matter how close between
two real numbers~$x$ and $y$, we can never guarantee $y\in\iota\left(x\right)$.

### 2017-09-26 prop_letter.tex
prop_letter.tex \lineskiplimit=1pt Added...
prop_letter.tex \section{Gleason} Updating...

\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{framed}
\newcommand{\amr}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{green}{Amr says: #1}\end{minipage}\end{framed}}
\newcommand{\yutsung}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{purple}{Yu-Tsung says: #1}\end{minipage}\end{framed}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}~\middle|~{#2}\right\} }}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\allowdisplaybreaks
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{\mathrm{L{#1}}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{\mathrm{R{#1}}}}}
\setcounter{secnumdepth}{3}
% https://tex.stackexchange.com/questions/345284/can-i-have-an-almost-non-breaking-space-in-latex
% Change non-breaking space to \nolinebreak[1] 
% Hence, it will not hyphenate too much surrounding words.
\newcommand{\rmd}{\mathrm{d}}
\newcommand{\rmi}{\mathrm{i}}

@@

Our framework allows us to generalize the above theorem to Born QIVPMs~$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$:
The collection of states consistent $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
gets closer and closer to $\rho$ as the intervals get narrower and
narrower, eventually collapsing to $\rho$ when the intervals become
infinitely precise real numbers covering the real line. To state this
result, we introduce a few technical notions. We call a state~$\rho$
a \emph{totally mixed state} if $0<\mu_{\rho}^{\mathrm{B}}\left(P\right)<1$
for any $P\in\events\backslash\left\{ \mathbb{0},\mathbb{1}\right\} $.
For example, $\rho_{0}$ and $\rho_{1}$ are both totally mixed states,
but any pure state is not totally mixed. We also define a notion of
distance between two states~$\rho$ and $\rho'$ as $\left\Vert \rho-\rho'\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$~\citep{544199},
and the set of all states whose distance with $\rho$ is bounded by
$\epsilon$ is denoted by $B_{\epsilon}\left(\rho\right)=\set{\rho'}{\left\Vert \rho-\rho'\right\Vert \le\epsilon}$.
By expending the definitions, we have $\rho'\in B_{\epsilon}\left(\rho\right)$
if and only if $\left|\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)\right|\le\epsilon$
for any pure state~$\ket{\psi}$. This fact will be used to prove
the following theorem that is obtained by replacing the QIVPM~$\bar{\mu}$
in Thm.~\ref{cor:Gleason's} by a Born QIVPM~$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
and estimate the states consistent with it.

\begin{thm}[Finite-precision Extension of Gleason's theorem for the
Born QIVPMs]\label{thm:Finite-precision-Gleason}In a Hilbert space
$\Hilb$ of dimension $3$, given any totally mixed state~$\rho$,
\begin{equation}
B_{\frac{1}{12n}}\left(\rho\right)\subseteq\coreBorn\left(\iota_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)\subseteq B_{1/n}\left(\rho\right)\,.
\end{equation}
\end{thm}

\noindent On the one hand, if a state~$\rho'$ is consistent with
$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$, the difference between the
Born quantum probabilities on any 1-dimensioal projector~$\proj{\psi}$,
$\left|\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)\right|$,
must be bounded by $1/n$ which is the maximal length of the intervals
in $\mathscr{I}_{n}$, i.e., $\rho'\in B_{1/n}\left(\rho\right)$.
On the other hand, if $\rho'\in B_{\frac{1}{12n}}\left(\rho\right)$,
we have $\left|\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)\right|\le\frac{1}{12n}$
for any pure state~$\ket{\psi}$. Since the dimension of the Hilbert
space is $3$, the same difference on any projector~$P$, $\left|\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)-\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)\right|$,
is bounded by $\frac{1}{4n}$. Moreover, the interiors of the intervals
in $\mathscr{I}_{n}$ shrinks $\frac{0.5}{2n}$ from the both ends,
excepts $\imposs$, $\left[0,\tfrac{1}{2n}\right]$, $\left[\tfrac{2n-1}{2n},1\right]$,
and $\necess$. Together with the fact that $0<\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)<1$
and $0<\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)<1$, $\mu_{\rho_{0}}^{\mathrm{B}}\left(P\right)$
and $\mu_{\rho_{1}}^{\mathrm{B}}\left(P\right)$ must both be in the
same intervals, and $\rho'$ is consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$.

As a concrete example of Thm.~\ref{thm:Finite-precision-Gleason},
since $\left\Vert \rho_{0}-\rho_{1}\right\Vert \le\frac{1}{24}$~\citep{HOSTunpublished},
we have $\rho_{0}\in B_{1/24}\left(\rho_{1}\right)$ and $\rho_{1}\in B_{1/24}\left(\rho_{0}\right)$
that should imply $\rho_{1}\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$
and $\rho_{0}\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}},\events\right)$,
and this is indeed the case for $P\in\events_{C}$ according to Table~\ref{tab:Compute-the-QIVPM}.
In general, as $n$ goes to infinity, only the state very close to
$\rho$ are guaranteed to belong to $\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$,
and the states consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
will become closer and closer to $\rho$.

>>>>

\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{framed}
\newcommand{\amr}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{green}{Amr says: #1}\end{minipage}\end{framed}}
\newcommand{\yutsung}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{purple}{Yu-Tsung says: #1}\end{minipage}\end{framed}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}~\middle|~{#2}\right\} }}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\allowdisplaybreaks
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{\mathrm{L{#1}}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{\mathrm{R{#1}}}}}
\setcounter{secnumdepth}{3}
% https://tex.stackexchange.com/questions/345284/can-i-have-an-almost-non-breaking-space-in-latex
% Change non-breaking space to \nolinebreak[1] 
% Hence, it will not hyphenate too much surrounding words.
\newcommand{\rmd}{\mathrm{d}}
\newcommand{\rmi}{\mathrm{i}}
\lineskiplimit=1pt

@@

Our framework allows us to generalize the above theorem to Born QIVPMs~$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$:
The collection of states consistent $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
gets closer and closer to $\rho$ as the intervals get narrower and
narrower, eventually collapsing to $\rho$ when the intervals become
infinitely precise real numbers covering the real line. To state this
result, we need to introduce a notion of distance between two states~$\rho$
and $\rho'$ as $\left\Vert \rho-\rho'\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$~\citep{544199}.
The following theorem that is obtained by replacing the QIVPM~$\bar{\mu}$
in Thm.~\ref{cor:Gleason's} by a Born QIVPM~$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
and estimate the states consistent with it.

\begin{thm}[Finite-precision Extension of Gleason's theorem for the
Born QIVPMs]\label{thm:Finite-precision-Gleason}In a Hilbert space
$\Hilb$ of dimension $3$, given any state~$\rho$, if a state~$\rho'\in\coreBorn\left(\iota_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)$,
then their distance~$\left\Vert \rho-\rho'\right\Vert \le\frac{1}{n}$.\end{thm}

\noindent If a state~$\rho'$ is consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
the difference between the Born quantum probabilities on any 1-dimensioal
projector~$\proj{\psi}$, $\left|\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)\right|$,
must be bounded by $1/n$ which implies $\left\Vert \rho-\rho'\right\Vert \le\frac{1}{n}$
by definition. As a concrete example of Thm.~\ref{thm:Finite-precision-Gleason},
since $\rho_{1}\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$
and $\rho_{0}\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}},\events\right)$
which is easy to see on the commuting subspace of event~$\events_{C}$
according to Table~\ref{tab:Compute-the-QIVPM}, we have $\left\Vert \rho_{0}-\rho_{1}\right\Vert \le\frac{1}{2}$,
and this is indeed the case because $\left\Vert \rho_{0}-\rho_{1}\right\Vert =\frac{1}{24}$~\citep{HOSTunpublished}.
